---
title: "Wasa Triathlon EDA App"
author: "Alejandro Coy"
date: '2019-03-04'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(lubridate)
library(stringr)
library(tidyr)
library(readxl)

#--------------------------------------------------------
wasaData18 <- read_excel("wasaData.xlsx", sheet = "2018")
wasaData17 <- read_excel("wasaData.xlsx", sheet = "2017")
wasaData16 <- read_excel("wasaData.xlsx", sheet = "2016")
wasaData15 <- read_excel("wasaData.xlsx", sheet = "2015")

wasaData18 <- rename(wasaData18, Division ="20")
cleandata <- function(x,y) {
  x %>%
    rename(category ="Division",age_group = "Place/Division",genderP ="Place/Gender")%>%
    separate(age_group,into = c("AgePosition","ToalAge"), sep = "/")%>%
    separate(genderP,into = c("GenderPosition","ToaGender"), sep = "/")%>%
    separate(Place,into = c("Position","TotalP"), sep = "/")%>%
    fill(Swim, T1,Bike,T2,Run)%>%
    select(-c("Event","Race #","ToalAge","ToaGender","TotalP"))%>%
    mutate_at(c(5:6,1),as.numeric)%>%
    filter(str_detect(category,"(^F|^M)(?=\\d)"))%>%
    mutate(category=as.factor(category))%>%
    mutate(gender = ifelse(str_detect(category,"F"),"F","M"))%>%
    mutate(year = as.factor(y))%>%
    mutate(category = str_replace(category,"[A-Z]",""))%>%
    filter(!is.na(Time))
}

wasaData18c <- cleandata(wasaData18,2018)
wasaData17c <- cleandata(wasaData17,2017)
wasaData16c <- cleandata(wasaData16,2016)
wasaData15c <- cleandata(wasaData15,2015)

wasaData <- bind_rows(wasaData18c,wasaData17c,wasaData16c,wasaData15c)
wasaData$year <- as.factor(wasaData$year) 
wasaDataF <- wasaData %>%
  filter(year == 2018)%>%
                filter((category=="3034" & gender=="F"))
yLabels <- function(x)
{
  x <- seconds_to_period(x)
  paste(hour(x),minute(x),second(x), sep = ':')
}
```

## Wasa Triathlon EDA App. 
In this exploratory data analysis project, we will explore the 2018 results for the Olympic race distance (1.5 km swim, 40 km bike ,10 km run). I retrieved the data from the official timing company’s websiteand a copy of the file is available in the GitHub folderof the project."

In its promotional campaign, the race is described as: “Gerick Sports Wasa Lake Triathlon features elite prize money and attracts some of the fastest triathletes in the West. Age groupers and elites compete together on a scenic and fast course.” Wasa Lake is located in the beautiful province of British Columbia just west from the Rocky Mountains and 4 hours away from Calgary.


              


## Use of the App

To check the reulst you have two tab option: Overall results and by Categories. 

After Selection you will have the option to filter the results by gender and year. You can select multiple years to compare


## Results 
An example  graph from the overall tab is time vs position in the race for all years. We could see how 2018 was the slowest of the last five years 
```{r Results, echo = TRUE,fig.width=4,fig.height=2.5,fig.align='center'}
  ggplot(data = wasaData, aes(x= Position, y = Time, color=year))+
            geom_point()+
            theme_light()
```

## Results
Another useful information is to be able to compare the time for each cactegory in each sport. In tab "Categories" we can compare between all the variable.

```{r echo =TRUE,fig.width=4,fig.height=2.5,fig.align='center'}
            ggplot(data = wasaDataF, aes( y = Run, x=AgePosition,color =year))+
              geom_line()+
              geom_point()
          
```

