---
title: 'Peer-graded Assignment: Statistical Inference Course ProjectPart 1'
author: "Alejandro Coy"
date: '2019-02-15'
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Peer-graded Assignment: Statistical Inference Course Project

The project consists of two parts:

1. A simulation exercise.
2. Basic inferential data analysis.

## Part 1: Simulation Excercise :

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

### Show the sample mean and compare it to the theoretical mean of the distribution.

Theoretical Mean of distribution is 1/ lambda if lmabda is 0.2, the theoretical mean is 5

```{r echo= TRUE}

set.seed(150)
lambda = 0.2
n = 40 #number of exponentials
n_simulations =1000

exp_mean = NULL
for(i in 1:n_simulations){
  exp_mean = c(exp_mean,mean(rexp(n,lambda)))
}
mean(exp_mean)

```

Result indicate that the mean of 40 exponentials for 10000 simulations is close to the theoretical mean of 5. 

### Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

The theoretical standart deviation is 1/lambda/sqrt(n) if, lamda is 0.2, the standart deviations is 5 and theoretical variance is 0.79
```{r echo = TRUE}
theo_variance <- (1/lambda/sqrt(n))^2
variance_exp <- var(exp_mean)
variance_exp
theo_variance
```

In this case the simulation is a little further.

### Show that the distribution is approximately normal.

To probe normality the exp_mean could be plotted in a histrogram

```{r echo=TRUE}
library(ggplot2)

data <- data.frame(exp_mean,n)
ggplot(data = data,aes(exp_mean))+
  geom_histogram(aes(y=..density..),fill = "steelBlue",binwidth = 0.2)+
  stat_function(fun = dnorm,args= list(mean=1/lambda,sd=1/lambda/sqrt(40)),color ="red")
```






