---
title: 'Peer-graded Assignment: Statistical Inference Course ProjectPart 2'
author: "Alejandro Coy"
date: '2019-02-15'
output:
  word_document: default
  html_document: default
---

# Part 2: Basic inferential data analysis. :

 The datased is imported and basic data information is obtained with the following functions: 

``` {r echo=TRUE}
library(datasets)
library(dplyr)
library(ggplot2)
data <- ToothGrowth
glimpse(data)
str(data)
head(data)
```
### Summary of the data:
Description

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

Usage

ToothGrowth
Format

A data frame with 60 observations on 3 variables.

[,1]	len	numeric	Tooth length
[,2]	supp	factor	Supplement type (VC or OJ).
[,3]	dose	numeric	Dose in milligrams/day

### Exploratory data analysis

```{r echo=TRUE}
ggplot(data,aes(factor(dose),y=len,fill=factor(dose)))+
  geom_boxplot()+
  facet_wrap(~supp)+
  theme_bw()+
  labs(x="Dose in mm",y="Tooth Lenght",fill="Suppy Method")
```

From the plot it can be seen the dose seems to have an effect in the tooth groth. However, it is not clear that the supplement type have an effect. 

### Hypothesis Testing

To compare the effect of the doses we will be using hypothesis testing between the three doses. First we have to filter the data by dose

```{r echo=TRUE}
dose_0.5 <- data %>%
  filter(dose == 0.5)
dose_1 <- data %>%
  filter(dose == 1)
dose_2 <- data %>%
  filter(dose == 2)
#t-tes Between doses
t.test(dose_0.5$len,dose_1$len,paired = FALSE)
t.test(dose_0.5$len,dose_2$len,paired = FALSE)
t.test(dose_1$len,dose_2$len,paired = FALSE)

```

From these results it can be infered that there is a significant difference between each on of the doses. The interval of confidence can be esay compute using conf 

```{r echo=TRUE}
t.test(dose_0.5$len,dose_1$len,paired = FALSE)$conf
t.test(dose_0.5$len,dose_2$len,paired = FALSE)$conf
t.test(dose_1$len,dose_2$len,paired = FALSE)$conf
```

Now we can compare betwen supplement type. Since we have stablished that the dose has a significant effect, the suplement type will be compare for each of doses and not between them. 

``` {r echo=TRUE}

vc_0.5 <- data %>%
  filter(dose == 0.5, supp =="VC")
vc_1 <- data %>%
  filter(dose == 1, supp =="VC")
vc_2 <- data %>%
  filter(dose == 2, supp =="VC")
oj_0.5 <- data %>%
  filter(dose == 0.5, supp =="OJ")
oj_1 <- data %>%
  filter(dose == 1, supp =="OJ")
oj_2 <- data %>%
  filter(dose == 2, supp =="OJ")
#Hypothesis testin
t.test(vc_0.5$len,oj_0.5$len,paired = FALSE)
t.test(vc_1$len,oj_1$len,paired = FALSE)
t.test(vc_2$len,oj_2$len,paired = FALSE)
# Confidence Intervals
t.test(vc_0.5$len,oj_0.5$len,paired = FALSE)$conf
t.test(vc_1$len,oj_1$len,paired = FALSE)$conf
t.test(vc_2$len,oj_2$len,paired = FALSE)$conf
```

From the result we can see that the supplmenet type has a signifcant effect for doses of 0.5 a 1 mg/day. In the case the 2 mg/day there is not change in the tooth growth by changing the supplment type. 

### Conclusions 

As conclusion the dose has a significant effect in the tooth growth. The confidence didn't contain 0 and they were negative. This indicate that increase of doses yield greater growth. 

In the case of the supplement type, the effect was just seen in the 0.5 and 1 mg.day dose. The intervals were negative, indicating that Orange Juice was a better delivery method. For the case of dose 2, there is not significant difference in the growth.