---
title: "ProjectRegressionModels"
author: "Alejandro Coy"
date: '2019-02-23'
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Peer-graded Assignment: Regression Models Course Project

###Executivy summary 

The main objective of this projet is to apply the regressions model to answer the following questions : 

1. “Is an automatic or manual transmission better for MPG”?
2. "Quantify the MPG difference between automatic and manual transmissions" ? 

Using exploratory data anlysis the difference between automatic and manual transmission cars was evident. Furthermore a t test was perfomed for the two groups showing a significant difference between the groups, showing the automatic cats achived less MPG than manual cars. 

For quantifiy the difference a linear regrtession model was used. Thre models were tested with differente explanatory variables.  With the best model it was determined that to have a manual transmission increase 1.63 MPG when all the other variable are held constant. 

### Exploratory Data Analysis

The dataset is preload in Rstudio. The columns with categorical value were changd to factor using the mutate function. 

```{r}
library(dplyr)
library(ggplot2) 
library(GGally)

data <- mtcars %>%
  mutate_at(c("cyl","gear","carb","am","vs"), as.factor)
levels(data$am) <- list(automatic="0", manual="1")
```

A serie of plots using ggpairs and gplot were done to see the relationship between MPG and the varaible (see APEENDIX)  

The first exploration indicates that manual cars yield higher MPG.

### T-Test

In order to perform  a t-test the data was divided depending on the type of transmission. Then the t.test function was performed:

```{r echo = FALSE}
automatic <- data %>%
  filter(am =="automatic" )
manual<- data %>%
  filter(am =="manual")
```
```{r }
t.test(automatic$mpg,manual$mpg,paired = FALSE)

```

As can be seen there is a significant diffrente with a lower mean for automatic cars.  The interval confidence does not contain 0 and p value is lowwer than 0.05

### Regression models

In order to quantify the difference between automatic and manual transmissions three linear models were used. 

#### Model 1 
Model 1 just consider the type of tranmission (am) as explanatory varaible

```{r echo=FALSE}
model1 <- lm(mpg ~ am, data = data)
```

The coefficient of the am variable indicates that the use of manual transmission increase 7.245 MPG, However, the R-squared is low indicating the change in transmission can just explain 35.9% the variation of the MPG. 

#### Model2

The model 2 take in consideration the variable that seem to have influence according to the pair plot from the EDA: 

```{r echo=FALSE}
model2 <- lm(mpg ~hp + drat + am + wt  , data= data)
```

The R-square for this model is 0.8428 with Adjusted R-squared : 0.8196. This indicates a better model. The residual plot are distributed around 0 which indicates a good fitting model. 

### Model3 

Finlaly model 3 take in considreation all the variables

```{r echo = TRUE}
model3 <- lm(mpg ~., data= data)
```

The R-square for this model is 0.8931, but an Adjusted R-squared : 0.779. This is an indication of overfitting of the model.  Finally we can compare between the models using the anova anlysis:
```{r echo = TRUE}
anova(model1,model2,model3)
```
From this results we can conclude there is not need to add all the variables in model 3 since there is not significant difference in the predections. However in comparison with model 1, model 2 does a significant better job (p-value lower than 0.05).

### APPENDIX

GRID PLOTS

```{r }
ggpairs(mtcars)
ggplot(data = mtcars, aes(y =am, x = mpg))+
  geom_point()
```

MEAN MPG FOR DIFFERENT TYPE OF TRANSMISSION

```{r echo-TRUE}
ggplot(data = data)+
  aes(x = am, y = mpg,fill=am)+
  geom_boxplot()
```

REISIDUAL PLOT FOR MODEL 2
```{R}
qplot(predict(model2), resid(model2))
```